---
title: "TS - ARIMAX"
author: "Swathi Ganesan"
date: "2023-08-10"
output: html_document
---

# ARIMA MODELING
```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


## Importing Libraries
```{r, message=FALSE, echo=FALSE}
# Load required packages
library(readxl)
library(ggplot2)
library(forecast)
library(ggseas)
library(zoo)
library(tseries)
library(fpp)
library(MASS)
library(ggplot2)

library(TSA)
```

## Loading Data
```{r}
# Read the Excel file
data <- read_excel("TS_Project_Data.xlsx")

# Convert the 'Period' column to year-quarter format
data$Period <- as.yearqtr(data$Period, format = "%Y Q%q")

# Convert "Period" to a Date object
data$Period <- as.Date(paste(data$Period, "01", sep = "-"), format = "%Y Q%m-%d")

# Extract year and quarter components
data$Year <- as.numeric(format(data$Period, "%Y"))
data$Quarter <- as.numeric(format(data$Period, "%m"))

# Create a time series object
ts_gdp <- ts(data$GDP, start = c(data$Year[1], data$Quarter[1]), frequency = 4)
ts_ipi <- ts(data$IPI, start = c(data$Year[1], data$Quarter[1]), frequency = 4)
ts_int <- ts(data$InterestRates_AssetPrices, start = c(data$Year[1], data$Quarter[1]), frequency = 4)
ts_cap <- ts(data$Labour_Capital_Prod, start = c(data$Year[1], data$Quarter[1]), frequency = 4)

# Create a matrix of exogenous variables
exog_matrix <- cbind(ts_ipi, ts_int, ts_cap)
exog_train <- window(exog_matrix, end = c(2010, 4))
exog_test <- window(exog_matrix, start = c(2011, 1))

#Plotting the time series
tsdisplay(ts_gdp, main = "UK GDP between 1955 and 2015")
```
## Assumption 1 - Constant Variance using BoxPlot Transformation

```{r}
# Load required packages
bc_lambda=BoxCox.lambda(ts_gdp)
ts_gdp %>% BoxCox(lambda = bc_lambda) %>% autoplot()
ts_gpd_bc = BoxCox(ts_gdp, lambda = bc_lambda)

# If the data show variation that increases or decreases with the level of the series, then a boxcox transformation can be useful.
```

## Assumption 2 - Stationarity is important for fitting Arima Models
ARIMA uses previous lags of series to model its behavior; therefore, modeling stable series with consistent properties involves less uncertainty.

```{r}
# RUNNING ADF AND KPSS TEST
adf_test <- adf.test(ts_gpd_bc)
print(adf_test)
kpss_test <- kpss.test(ts_gpd_bc)
print(kpss_test)

```

Data can be seen as non-stationary

```{r}
# Differencing the time series
ts_gdp_bc_diff <- diff(ts_gpd_bc)

# Check the stationarity of the differenced series using ADF and KPSS tests
adf_test_result_diff <- adf.test(ts_gdp_bc_diff)
kpss_test_result_diff <- kpss.test(ts_gdp_bc_diff)

print(adf_test_result_diff)
print(kpss_test_result_diff)

tsdisplay(ts_gdp_bc_diff)
```

Data seems stationary post single order differencing.
From observations, we can see an exponential decay in the ACF plot, along with hard drops in the PACF plot.
This indicates that the process might be an Autoregressive process of Order 3 

Since both ACF and PACF are trailing off, this model can also be an ARMA model 

### Seasonal differencing is needed for Seasonal Arima
```{r}
ts_gdp_train_seasonal_diff <- diff(ts_gdp_train, difference = 4)
tsdisplay(ts_gdp_train_seasonal_diff)
``` 

## Dividing into Train and Test prior to Model Fitting
```{r}
# Perform train-test split
ts_gdp_train <- window(ts_gdp, end = c(2010, 4))  # Training data until 2009 Q4
ts_gdp_test <- window(ts_gdp, start = c(2011, 1)) # Test data from 2010 Q1 onwards
number_of_forecasts = 20

# Plot time series of train and test data
ggplot() +
  geom_line(data = as.data.frame(ts_gdp_train), aes(x = time(ts_gdp_train), y = ts_gdp_train), color = "blue", linetype = "solid", size = 1) +
  geom_line(data = as.data.frame(ts_gdp_test), aes(x = time(ts_gdp_test), y = ts_gdp_test), color = "red", linetype = "solid", size = 1) +
  labs(title = "Time Series of Train and Test Data", x = "Year", y = "Value") +
  theme_minimal()
```

## Model 1 - Non Seasonal Auto ARIMAX
```{r}
arimax_model <- auto.arima(ts_gdp_train, xreg =window(ts_ipi, end = c(2010, 4)), seasonal = FALSE, lambda = bc_lambda, trace = TRUE)
arimax_model
eacf(ts_gdp_train)
```


## Analysing Forecasts from the Non-Seasonal ARIMAX Model
```{r}
# Forecast using the ARIMAX model
forecast_arimax <- forecast(arimax_model, xreg = window(ts_ipi, start = c(2011, 1)), h = number_of_forecasts)

# Plot the forecast results
plot(forecast_arimax)

# Checking for RMSE
squared_diff_arimax <- (forecast_arimax$mean - ts_gdp_test)^2
rmse_arimax <- sqrt(mean(squared_diff_arimax))

# Calculate accuracy metrics (e.g., MAE, RMSE) to evaluate the forecast
accuracy(forecast_arimax, ts_gdp_test)

# Check residuals
checkresiduals(residuals(forecast_arimax))

# Print RMSE
print(paste("RMSE:", rmse_arimax))
```
#ARIMA(3,1,2) 
The Ljung-Box test evaluates the null hypothesis that the residuals are not autocorrelated. In this case, the high p-value of 0.9527 suggests that there is no significant evidence to reject the null hypothesis. This indicates that the residuals are likely behaving as white noise, showing no substantial autocorrelation at the specified lags.


## Model 2 - Seasonal ARIMAX
```{r}
arimax_model_seasonal <- auto.arima(ts_gdp_train, xreg =window(ts_ipi, end = c(2010, 4)), seasonal = TRUE, lambda = bc_lambda, trace = TRUE)

arimax_model_seasonal
```

## Analysing Forecasts from the Seasonal ARIMAX Model

```{r}
# Forecast using the ARIMAX model
forecast_arimax <- forecast(arimax_model_seasonal, xreg = window(ts_ipi, start = c(2011, 1)), h = number_of_forecasts)

# Plot the forecast results
plot(forecast_arimax)

# Checking for RMSE
squared_diff_arimax <- (forecast_arimax$mean - ts_gdp_test)^2
rmse_arimax <- sqrt(mean(squared_diff_arimax))

# Calculate accuracy metrics (e.g., MAE, RMSE) to evaluate the forecast
accuracy(forecast_arimax, ts_gdp_test)

# Check residuals
checkresiduals(residuals(forecast_arimax))

# Print RMSE
print(paste("RMSE:", rmse_arimax))
```

## Model 3 - Multiple exogenoeus variables with non-seasonal ARIMAX
```{r}
arimax_model_exog <- auto.arima(ts_gdp_train, xreg =exog_train, seasonal = FALSE, lambda = bc_lambda, trace = TRUE)

arimax_model_exog
```

```{r}
# Forecast using the ARIMAX model
forecast_arimax <- forecast(arimax_model_exog, xreg = exog_test, h = number_of_forecasts)

# Plot the forecast results
plot(forecast_arimax)

# Checking for RMSE
squared_diff_arimax <- (forecast_arimax$mean - ts_gdp_test)^2
rmse_arimax <- sqrt(mean(squared_diff_arimax))

# Calculate accuracy metrics (e.g., MAE, RMSE) to evaluate the forecast
accuracy(forecast_arimax, ts_gdp_test)

# Check residuals
checkresiduals(residuals(forecast_arimax))

# Print RMSE
print(paste("RMSE:", rmse_arimax))
```

## Model 4 - Multiple exogenoeus variables with seasonal ARIMAX
```{r}
arimax_model_seasonal_exog <- auto.arima(ts_gdp_train, xreg =exog_train, seasonal = TRUE, lambda = bc_lambda, trace = TRUE)

arimax_model_seasonal_exog
```

```{r}
# Forecast using the ARIMAX model
forecast_arimax <- forecast(arimax_model_seasonal_exog, xreg = exog_test, h = number_of_forecasts)

# Plot the forecast results
plot(forecast_arimax)

# Checking for RMSE
squared_diff_arimax <- (forecast_arimax$mean - ts_gdp_test)^2
rmse_arimax <- sqrt(mean(squared_diff_arimax))

# Calculate accuracy metrics (e.g., MAE, RMSE) to evaluate the forecast
accuracy(forecast_arimax, ts_gdp_test)

# Check residuals
checkresiduals(residuals(forecast_arimax))

# Print RMSE
print(paste("RMSE:", rmse_arimax))
```


## Model 5: VAR Model
#VAR models are used to capture the relationships and interactions among multiple time series variables.

```{r}
# Check out other models
VARselect(cbind(log(data$GDP), log(data$IPI)), lag.max = 10, type="both", season = 12)$selection
```

We can proceed by picking the lag order based on the AIC (Akaike Information Criterion) for the ability to capture the underlying patterns in the data with a simple method and hence proceed with p=5.

```{r}
# Assuming df is your data frame
train_rows <- data[1:224, c("GDP", "IPI")]
var_gdp_ipi <- VAR(cbind(log(train_rows$GDP), log(train_rows$IPI)), p = 5, type = "both", season = 4)

pred <- predict(var_gdp_ipi, n.ahead = number_of_forecasts)

# Extract predicted GDP starting from a specific period
gdp_pred <- ts(pred$fcst$y1[, "fcst"], start = c(2011, 1), frequency = 4)
ipi_pred <- ts(pred$fcst$y2[, "fcst"], start = c(2011, 1), frequency = 4)

# Plot predicted GDP
ts.plot(cbind(log(ts_gdp_train), gdp_pred), col = c("blue", "red"),
        lty = c(1, 1),
        lwd = c(1, 2), main = "Forecast of GDP", ylab = "log(GDP)", xlab = "Period") 

# Plot predicted GDP
ts.plot(cbind(window(log(ts_ipi), end = c(2010, 4)), ipi_pred), col = c("blue", "red"),
        lty = c(1, 1),
        lwd = c(1, 2), main = "Forecast of IPI", ylab = "log(IPI)", xlab = "Period") 
```

```{r}
# Checking for RMSE
squared_diff <- (exp(pred$fcst$y1[, "fcst"]) - ts_gdp_test)^2
rmse_var <- sqrt(mean(squared_diff))

# Print RMSE
print(paste("RMSE:", rmse_var))
```
