---
title: "Time Series Final Project - Intervention Modeling"
author: "Kshitij Mittal"
date: "2023-08-08"
output: html_document
---

# Intervention Analysis
Time series are frequently affected by certain external events such as holidays, strikes, sales promotions, and other policy changes. We call these external events interventions.
It is assumed that the intervention affects the process by changing the mean function or trend of a time series.

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


## Importing Libraries
```{r, message=FALSE, echo=FALSE}
# Load required packages
library(readxl)
library(ggplot2)
library(forecast)
library(ggseas)
library(zoo)
library(tseries)
library(fpp)
library(MASS)
library(ggplot2)
library(lmtest)
library(strucchange)

library(TSA)
```

## Loading Data
```{r}
# Read the Excel file
data <- read_excel("TS_Project_Data.xlsx")

# Convert the 'Period' column to year-quarter format
data$Period <- as.yearqtr(data$Period, format = "%Y Q%q")

# Convert "Period" to a Date object
data$Period <- as.Date(paste(data$Period, "01", sep = "-"), format = "%Y Q%m-%d")

# Extract year and quarter components
data$Year <- as.numeric(format(data$Period, "%Y"))
data$Quarter <- as.numeric(format(data$Period, "%m"))

# Create a time series object
ts_gdp <- ts(data$GDP, start = c(data$Year[1], data$Quarter[1]), frequency = 4)

#Plotting the time series
autoplot(ts_gdp, main = "UK GDP between 1955 and 2015")
tsdisplay(ts_gdp, main = "UK GDP between 1955 and 2015")
```

From the above plot, we can see a few sudden dips in the model. These might be intervention effects. 
Few economic downturns observed in this time frame:

- 1973 oil crisis
- Early 1980s recession
- Early 1990s recession - US Savings and Loan Crisis
- Great Recession - 2008 Financial Crisis

https://en.wikipedia.org/wiki/List_of_recessions_in_the_United_Kingdom

## Intervention Analysis Theory
Yt = mt + Nt
We can model the change in the mean function by an ARMA-type specification


## Assumption 1 - Constant Variance using BoxPlot Transformation

```{r}
# Load required packages
bc_lambda=BoxCox.lambda(ts_gdp)
ts_gdp %>% BoxCox(lambda = bc_lambda) %>% autoplot()
ts_gpd_bc = BoxCox(ts_gdp, lambda = bc_lambda)


# If the data show variation that increases or decreases with the level of the series, then a boxcox transformation can be useful.
```

## Assumption 2 - Stationarity is important for fitting Arima Models
ARIMA uses previous lags of series to model its behavior; therefore, modeling stable series with consistent properties involves less uncertainty.

```{r}
# RUNNING ADF AND KPISS TEST
adf_test <- adf.test(ts_gpd_bc)
print(adf_test)
kpss_test <- kpss.test(ts_gpd_bc)
print(kpss_test)

```

Data can be seen as non-stationary

```{r}
# Differencing the time series
ts_gdp_bc_diff <- diff(ts_gpd_bc)

# Check the stationarity of the differenced series using ADF and KPSS tests
adf_test_result_diff <- adf.test(ts_gdp_bc_diff)
kpss_test_result_diff <- kpss.test(ts_gdp_bc_diff)

print(adf_test_result_diff)
print(kpss_test_result_diff)

tsdisplay(ts_gdp_bc_diff)
```

Data seems stationary post single order differencing.
From observations, we can see an exponential decay in the ACF plot, along with hard drops in the PACF plot.
This indicates that the process might be an Autoregressive process of Order 3 

Since both ACF and PACF are trailing off, this model can also be an ARMA model 

## Finding Train vs Test Data

```{r}
ts_gdp_train <- window(ts_gdp, end = c(2010, 4))  # Training data until 2010 Q4
ts_gdp_test <- window(ts_gdp, start = c(2011, 1)) # Test data from 2011 Q1 onwards
number_of_forecasts = 20

ts_gdp_train_pre_int = window(ts_gdp_train, end = c(2008, 1))
ts_gdp_train_post_int = window(ts_gdp_train, start = c(2008, 2))

# Plot time series of train and test data
ggplot() +
  geom_line(data = as.data.frame(ts_gdp_train), aes(x = time(ts_gdp_train), y = ts_gdp_train), color = "blue", linetype = "solid", size = 1) +
  geom_line(data = as.data.frame(ts_gdp_test), aes(x = time(ts_gdp_test), y = ts_gdp_test), color = "red", linetype = "solid", size = 1) +
  labs(title = "Time Series of Train and Test Data", x = "Year", y = "Value") +
  theme_minimal()

ggplot() +
  geom_line(data = as.data.frame(ts_gdp_train_pre_int), aes(x = time(ts_gdp_train_pre_int), y = ts_gdp_train_pre_int), color = "blue", linetype = "solid", size = 1) +
  geom_line(data = as.data.frame(ts_gdp_train_post_int), aes(x = time(ts_gdp_train_post_int), y = ts_gdp_train_post_int), color = "red", linetype = "solid", size = 1) +
  labs(title = "Time Series of Train Data - Pre and Post Intervention", x = "Year", y = "Value") +
  theme_minimal()


```

## ARIMA Model for Pre-Intervention Data

```{r}
fit_auto_arima_pre_int=auto.arima(ts_gdp_train_pre_int, seasonal=FALSE, lambda = bc_lambda, trace = TRUE)
coeftest(fit_auto_arima_pre_int)
```

Auto Arima is giving a best model of ARIMA(0,1,3) - indicating the absence of an auto-regression component and only moving average.
Differencing (Integration) of order 1 is consistent with our prior observations.

## Intervention Modeling - Step Function

```{r}

# Assigning the breakpoint
breakpoint=215 # Index for 2008-Q2 in ts_gdp_train

# ASSIGNING A STEP FUNCTION FOR LEVEL
level <- c(rep(0, breakpoint), 
              rep(1, length(ts_gdp_train) - breakpoint))

fit_ar_int_1= Arima(ts_gdp_train, order = c(0,1,3), xreg = level, lambda = bc_lambda, include.drift = TRUE)
summary(fit_ar_int_1)
coeftest(fit_ar_int_1)

```

When we ran intervention modeling, our model fit improved improved from a non-intervention ARIMA(0,1,3) - Had AIC of 496, now we have 490

## Analysing Forecasts from the Non-Seasonal Arima Model
```{r}
forecast_arima_int_1=forecast(fit_ar_int_1, xreg = rep(1, number_of_forecasts), h= number_of_forecasts)
plot(forecast_arima_int_1, main = "Arima(0,1,3) + Intervention Modeling - Forecasted Values")

# Check residuals
checkresiduals(residuals(forecast_arima_int_1))
```
The residuals look normally distributed with no siginficant auto-correlation at any lags.
The model also passes the Ljung Box test, and we cannot reject the null hypothesis that there is no auto-correlation in the residuals.

```{r}
# Checking for RMSE
squared_diff_arima_int_1 <- (forecast_arima_int_1$mean - ts_gdp_test)^2
rmse_arima_int_1 <- sqrt(mean(squared_diff_arima_int_1))

# Print RMSE
print(paste("RMSE:", rmse_arima_int_1))
accuracy(forecast_arima_int_1, ts_gdp_test)
```


## Intervention Modeling - Step and Impulse using ARIMA
```{r}

# Assigning the breakpoint
breakpoint=215 # Index for 2008-Q2 in ts_gdp_train

# ASSIGNING A STEP FUNCTION FOR LEVEL
I2008F=1*(seq(ts_gdp_train)==215)
S2008F=1*(seq(ts_gdp_train)>=215)

fit_ar_int_2= Arima(ts_gdp_train, order = c(0,1,3), xreg = cbind(I2008F, S2008F), lambda = bc_lambda, include.drift = TRUE)
summary(fit_ar_int_2)
coeftest(fit_ar_int_2)

```

```{r}
# Checking for RMSE
forecast_arima_int_2=forecast(fit_ar_int_2, xreg = cbind(rep(0, number_of_forecasts),rep(1, number_of_forecasts)), h= number_of_forecasts)
squared_diff_arima_int_2 <- (forecast_arima_int_2$mean - ts_gdp_test)^2
rmse_arima_int_2 <- sqrt(mean(squared_diff_arima_int_2))

# Print RMSE
plot(forecast_arima_int_2)
accuracy(forecast_arima_int_2, ts_gdp_test)
print(paste("RMSE:", rmse_arima_int_2))

```

## Intervention Modeling - Step Function using ARIMAX

```{r}

# Assigning the breakpoint
breakpoint=215 # Index for 2008-Q2 in ts_gdp_train

# ASSIGNING A STEP FUNCTION FOR LEVEL
level <- c(rep(0, breakpoint), 
              rep(1, length(ts_gdp_train) - breakpoint))


fit_arimax <- arimax(BoxCox(ts_gdp_train, lambda = bc_lambda), order = c(0, 1, 3), xreg = level, method="ML")

# fit_arimax <- arimax(BoxCox(ts_gdp_train, lambda = bc_lambda), order = c(0, 1, 3), xtransf = matrix(level, ncol = 1), include.mean = TRUE, transfer=list(c(1,0)), method="ML")

# Summary of the ARIMAX model
summary(fit_arimax)
coeftest(fit_arimax)

```


## EXPERIMENT

```{r}

library(fpp2)
library(ggplot2)
library(TSA)

I2008F=1*(seq(ts_gdp_train)==215)
S2008F=1*(seq(ts_gdp_train)>=215)

autoplot(ts_gdp_train)
fit_arimax_2 = arimax(BoxCox(ts_gdp_train, lambda = bc_lambda),
                    order = c(0,1,3),
                    xtransf = data.frame(I2008F, S2008F),
                    transfer = list(c(0,0),c(1,0)),
                    method='ML',
                    include.mean = TRUE)

summary(fit_arimax_2)
coeftest(fit_arimax_2)
checkresiduals(residuals(fit_arimax_2))

```

```{r}

library(fpp2)
library(ggplot2)
library(TSA)

data("airmiles")
I911=1*(seq(ts_gdp_train)==69)
S911=1*(seq(ts_gdp_train)>=69)

autoplot(airmiles)
air.mPulse = arimax(log(airmiles),
                    order = c(0,1,1),
                    xtransf = data.frame(S911, I911),
                    transfer = list(c(0,0),c(1,0)),
                    xreg = data.frame(Dec96=1*(seq(airmiles)==12),
                                      Jan97=1*(seq(airmiles)==13),
                                      Dec02=1*(seq(airmiles)==84)),
                                      method='ML',
                    include.mean = TRUE)

air.mPulse              

```