---
title: "kshitji"
author: "Jeniffer Lee"
date: "2023-08-08"
output: html_document
---

# ARIMA MODELING
```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


## Importing Libraries
```{r, message=FALSE, echo=FALSE}
# Load required packages
library(readxl)
library(ggplot2)
library(forecast)
library(ggseas)
library(zoo)
library(tseries)
library(fpp)
library(MASS)
library(ggplot2)

library(TSA)
```

## Loading Data
```{r}

# Read the Excel file
data <- read_excel("TS_Project_Data.xlsx")

# Convert the 'Period' column to year-quarter format
data$Period <- as.yearqtr(data$Period, format = "%Y Q%q")

# Convert "Period" to a Date object
data$Period <- as.Date(paste(data$Period, "01", sep = "-"), format = "%Y Q%m-%d")

# Extract year and quarter components
data$Year <- as.numeric(format(data$Period, "%Y"))
data$Quarter <- as.numeric(format(data$Period, "%m"))

# Create a time series object
ts_gdp <- ts(data$GDP, start = c(data$Year[1], data$Quarter[1]), frequency = 4)


# Read the Excel file
data17 <- read_excel("tsgdp1700.xlsx", col_names = TRUE)
# Assuming your data frame is named 'data'
new_column_names <- c("year", "UKGDP")
# Rename the second column
names(data17)[2] <- new_column_names[2]

```
# Plot the Time Series
```{r}
#Plotting the time series
#plot(data17, ylab = "UK GDP", main = "Annual UK GDP (1700 - 2015)") 
#tsdisplay(data17, main = "Annual UK GDP (1700 - 2015)")

options(repr.plot.width = 20, repr.plot.height = 80)  # Adjust the values as needed
ggplot(data17, aes(x = year, y = UKGDP)) +
  geom_line() +
  labs(x = "Year", y = " UK nominal GDP (£mn)",
       title = "United Kingdom GDP" ,
      subtitle = "From 1700 to 2015") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels
  scale_x_continuous(breaks = seq(min(data17$year), max(data17$year), by = 50))  # Set x-axis breaks every 20 years
```

# Plot the Time Series
```{r}
#Plotting the time series
plot(ts_gdp, ylab = "UK GDP", main = "Quarterly UK GDP (1955 - 2015)") 

tsdisplay(ts_gdp, main = "UK GDP between 1955 and 2015")
```
As the above, ACF is not tailing off. So, we need to make the data the stationary data before building a model by considering BOXCOX transformation or differencing. 




```{r}
# RUNNING ADF AND KPSS TEST
adf_test <- adf.test(ts_gdp)
print(adf_test)
kpss_test <- kpss.test(ts_gdp)
print(kpss_test)

```

ADF test's P-value: 0.61 > 0.05 --> fail to reject the Null(time series has a unitroot and non-stationary)
KPSS test's P-value: 0.01 < 0.05 --> reject the Null(the timeseries is stationary).  

Data itself can be seen as non-stationary



## Assumption 1 - Constant Variance using BoxPlot Transformation

# If the data show variation that increases or decreases with the level of the series, then a boxcox transformation can be useful.
```{r}
par(mfrow = c(4, 3))
autoplot(ts_gdp)
ts_gdp %>% BoxCox(lambda = -1) %>% autoplot()
ts_gdp %>% BoxCox(lambda = -0.5) %>% autoplot()
ts_gdp %>% BoxCox(lambda = 0.0) %>% autoplot()
ts_gdp %>% BoxCox(lambda = 0.5) %>% autoplot()
ts_gdp %>% BoxCox(lambda = 1) %>% autoplot()
bc_lambda <-BoxCox.lambda(ts_gdp)
bc_lambda
ts_gdp %>% BoxCox(lambda = bc_lambda) %>% autoplot()
ts_gdp_bc = BoxCox(ts_gdp, lambda = bc_lambda)
```

ARIMA uses previous lags of series to model its behavior; therefore, modeling stable series with consistent properties involves less uncertainty.
```{r}
tsdisplay(ts_gpd_bc)
```

```{r}
# RUNNING ADF AND KPSS TEST
adf_test <- adf.test(ts_gpd_bc)
print(adf_test)
kpss_test <- kpss.test(ts_gpd_bc)
print(kpss_test)

```
ADF test's P-value: 0.11 > 0.05 --> fail to reject the Null(time series has a unitroot and non-stationary)
KPSS test's P-value: 0.01 < 0.05 --> reject the Null(the timeseries is stationary).  

Data adjusted boxcox transformation can be seen as non-stationary yet!



### what about the first order of differencing with boxcox transformed GDP.

```{r}
# Differencing the time series
ts_gdp_bc_diff <- diff(ts_gpd_bc)

# Check the stationarity of the differenced series using ADF and KPSS tests
adf_test_result_diff <- adf.test(ts_gdp_bc_diff)
kpss_test_result_diff <- kpss.test(ts_gdp_bc_diff)

print(adf_test_result_diff)
print(kpss_test_result_diff)

tsdisplay(ts_gdp_bc_diff)
```

Data seems stationary post single order differencing.
From observations, we can see an exponential decay in the ACF plot, along with hard drops in the PACF plot.
This indicates that the process might be an Autoregressive process of Order 3 

Since both ACF and PACF are trailing off, this model can also be an ARMA model 

## Dividing into Train and Test prior to Model Fitting
```{r}
# Perform train-test split
ts_gdp_train <- window(ts_gdp, end = c(2010, 4))  # Training data until 2009 Q4
ts_gdp_test <- window(ts_gdp, start = c(2011, 1)) # Test data from 2010 Q1 onwards
number_of_forecasts = 20

# Plot time series of train and test data
ggplot() +
  geom_line(data = as.data.frame(ts_gdp_train), aes(x = time(ts_gdp_train), y = ts_gdp_train), color = "blue", linetype = "solid", size = 1) +
  geom_line(data = as.data.frame(ts_gdp_test), aes(x = time(ts_gdp_test), y = ts_gdp_test), color = "red", linetype = "solid", size = 1) +
  labs(title = "Time Series of Train and Test Data", x = "Year", y = "Value") +
  theme_minimal()
```

## Model Fitting - Non Seasonal Auto Arima

```{r}
auto.arima(ts_gdp_train,seasonal=FALSE, lambda = bc_lambda, trace = TRUE)

# RUNNING EACF ON FIRST ORDER DIFFERENCE TRAIN DATA
eacf(diff(ts_gdp_train))
```

Auto Arima is giving a best model of ARIMA(0,1,3) - indicating the absence of an auto-regression component and only moving average.
Differencing (Integration) of order 1 is consistent with our prior observations.

EACF run for first order differenced data also indicates using (1,1) and (1,3) models. We can run these also to check if we observe better model fits


```{r}
fit_ar1=Arima(ts_gdp_train, order=c(0,1,3), include.drift = TRUE, lambda = bc_lambda) 
fit_ar2=Arima(ts_gdp_train, order=c(1,1,1), include.drift = TRUE, lambda = bc_lambda) 
fit_ar3=Arima(ts_gdp_train, order=c(1,1,3), include.drift = TRUE, lambda = bc_lambda) 

fit_ar1$aicc
fit_ar2$aicc
fit_ar3$aicc
```

We are still getting the Auto Arima indicated ARIMA(0,1,3) model with drift.
We will now check forecasts from this model

## Analysing Forecasts from the Non-Seasonal Arima Model
```{r}
forecast_arima=forecast(fit_ar1, h= number_of_forecasts)
plot(forecast_arima, main = "Arima(0,1,3) Forecasted Values")

# Check residuals
checkresiduals(residuals(forecast_arima))
```
The residuals look normally distributed with no siginficant auto-correlation at any lags.
The model also passes the Ljung Box test, and we cannot reject the null hypothesis that there is no auto-correlation in the residuals.

```{r}
# Checking for RMSE
squared_diff_arima <- (forecast_arima$mean - ts_gdp_test)^2
rmse_arima <- sqrt(mean(squared_diff_arima))

# Print RMSE
print(paste("RMSE:", rmse_arima))
```



# SARIMA Modeling 


As we can see the results of decomposition, even though it is hard to verify seasonality of the data with our own eyes, there seems to be a seasonality. So, seasonal Arima model could suggest a better model to explain the UK GDP. 

```{r}
plot(decompose(ts_gdp))
```


### Seasonal differencing is needed for Seasonal Arima
As we are going to consider the seasonality by transforming process, 
```{r}

# boxcox
ts_gdp_bc = BoxCox(ts_gdp, lambda = bc_lambda)

# FIRST: SEASONAL DIFFERENCE 
ts_gdp_bc_sdiff <- diff(ts_gdp_bc, difference = 4) 
tsdisplay(ts_gdp_bc_sdiff)

# Check the stationarity of the seasonally differenced series using ADF and KPSS tests
adf.test(ts_gdp_bc_sdiff)
kpss.test(ts_gdp_bc_sdiff)

``` 
From plot we see significant spikes at lag 1 and 2 in ACF

```{r}

# SECOND: Regular(NON-SEASONAL) DIFFERENCE 
ts_gdp_bc_sdiff_diff <- diff(ts_gdp_bc_sdiff) 
tsdisplay(ts_gdp_bc_sdiff_diff)

# Check the stationarity of the differenced series using ADF and KPSS tests
adf.test(ts_gdp_bc_sdiff_diff)
kpss.test(ts_gdp_bc_sdiff_diff)

``` 
ACF and PACF in Seasonal difference and First difference are showing similar with the just seasonal differenced ones.
These double differences might not be neccessary to apply. 


## Model Fitting 2 - Seasonal Auto Arima
```{r}
Model_sarima <-auto.arima(ts_gdp_train ,seasonal=TRUE, lambda = bc_lambda, trace = FALSE)
Model_sarima
coeftest(Model_sarima)

```

```{r}
fit_sar1=Arima(ts_gdp_train, order=c(2,1,2), seasonal = list(order = c(0,0,2),period=4), include.drift = TRUE, lambda = bc_lambda) 
fit_sar1$aicc
```

```{r}
checkresiduals(fit_sar1)
```

Jung test shows that p-value is 0.27 > 0.05 --> fail to reject(accept) the null hypothesis(the Timeseries is independently distributed)
The residuals look normally distributed with no significant auto-correlation at any lags.
The model also passes the Ljung Box test, and we cannot reject the null hypothesis that there is no auto-correlation in the residuals.



## Analysing Forecasts from the Seasonal Arima Model
```{r}
forecast_sarima=forecast(fit_sar1, h= number_of_forecasts)
plot(forecast_sarima, main = "SARIMA(2,1,2)(0,0,2)[4] Forecasted Values")

```


```{r}
# Checking for RMSE

squared_diff_sarima <- (forecast_sarima$mean - ts_gdp_test)^2
rmse_sarima <- sqrt(mean(squared_diff_sarima))

# Print RMSE
print(paste("RMSE:", rmse_sarima))
accuracy(forecast_sarima, ts_gdp_test)

```

## Plot the forecast
```{r}
autoplot(ts_gdp_train) +
  autolayer(forecast_sarima$mean, series="SARIMA")+
  ggtitle("Forecasts for quarterly UK GDP(2011-2015)") +
  xlab("time") + ylab("GDP")

```







### APPENDIX Transformation 

```{r}

# pre-intervention period

ly <- log(ts_gdp)
dly1 <-diff(ly)
dly4 <- diff(ly, lag=4)
d2ly1_4 <- diff(diff(ly), lag=4)

library(forecast)
# Original and transformed data
par(mfrow = c(2, 3))
plot(ts_gdp, main = expression(ts_gdp))
plot(ly, main = expression(log(ts_gdp)))

plot.new()
plot(dly1, ylab='', xlab='year', main=expression(paste(Delta, "log(y)")))
plot(dly4, ylab='', xlab='year', main=expression(paste(Delta[4], "log(y)")))
plot(d2ly1_4, ylab='', xlab='year', main=expression(paste(Delta,Delta[4], "log(y)")))


# ACFs and PACFs
maxlag <- 60
par(mfrow=c(2,4), mar=c(3,3,4,2))
Acf(ly, type="correlation", lag.max=maxlag, main=expression(paste("ACF for log(y)")))
Acf(dly1, type="correlation", lag.max=maxlag, main=expression(paste("ACF for ",Delta, "log(y)")))
Acf(dly4, type="correlation", lag.max=maxlag, main=expression(paste("ACF for ",Delta[4], "log(y)")))
Acf(d2ly1_4, type="correlation", lag.max=maxlag, main=expression(paste("ACF for ",Delta,Delta[4], "log(y)")))

Acf(ly, type="partial", lag.max=maxlag, main=expression(paste("PACF for log(y)")))
Acf(dly1, type="partial", lag.max=maxlag, main=expression(paste("PACF for ",Delta, "log(y)")))
Acf(dly4, type="partial", lag.max=maxlag, main=expression(paste("PACF for ",Delta[4], "log(y)")))
Acf(d2ly1_4, type="partial", lag.max=maxlag, main=expression(paste("PACF for ",Delta,Delta[4], "log(y)")))




```

```{r}

# pre-intervention period

dy1 <-diff(ts_gdp)
dy4 <- diff(ts_gdp, lag=4)
d2y1_4 <- diff(diff(ts_gdp), lag=4)


library(forecast)
# Original and transformed data
par(mfrow = c(2, 3))
plot(ts_gdp, main = expression(ts_gdp))

plot.new()
plot.new()
plot(dy1, ylab='', xlab='year', main=expression(paste(Delta, "y")))
plot(dy4, ylab='', xlab='year', main=expression(paste(Delta[4], "y")))
plot(d2y1_4, ylab='', xlab='year', main=expression(paste(Delta,Delta[4], "y")))


# ACFs and PACFs
maxlag <- 60
par(mfrow=c(2,4), mar=c(3,3,4,2))
Acf(ts_gdp, type="correlation", lag.max=maxlag, main=expression(paste("ACF for y")))
Acf(dy1, type="correlation", lag.max=maxlag, main=expression(paste("ACF for ",Delta, "y")))
Acf(dy4, type="correlation", lag.max=maxlag, main=expression(paste("ACF for ",Delta[4], "y")))
Acf(d2y1_4, type="correlation", lag.max=maxlag, main=expression(paste("ACF for ",Delta,Delta[4], "y")))

Acf(ts_gdp, type="partial", lag.max=maxlag, main=expression(paste("PACF for y")))
Acf(dy1, type="partial", lag.max=maxlag, main=expression(paste("PACF for ",Delta, "y")))
Acf(dy4, type="partial", lag.max=maxlag, main=expression(paste("PACF for ",Delta[4], "y")))
Acf(d2y1_4, type="partial", lag.max=maxlag, main=expression(paste("PACF for ",Delta,Delta[4], "y")))




```



